---
title: "ALK Construction - Exercise"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot[RO,RE]{Ogle (Nunavut, 6-9 Aug 2016)}
- \fancyfoot[LE,LO]{ALK Construction Exercise}
output: pdf_document
geometry: left=0.5in, right=0.5in, top=0.5in, bottom=0.75in
---

```{r echo=FALSE, results='hide', message=FALSE}
source("../knitr_setup.R")
showkey <- FALSE
opts_chunk$set(echo=showkey,
               results=ifelse(showkey,"markup","hide"),
               fig.show=ifelse(showkey,"asis","hide"))
```
```{r message=FALSE, warning=FALSE}
library(FSA)
library(FSAdata)
library(dplyr)
library(nnet)
```

[Wolfert (1980)](https://pubs.er.usgs.gov/publication/1000461) measured the total length (TL) of 1288 [Rock Bass](https://en.wikipedia.org/wiki/Rock_bass) (*Ambloplites rupestris*) from Eastern [Lake Ontario](https://en.wikipedia.org/wiki/Lake_Ontario) in the late 1970s.  In addition, [scales](https://en.wikipedia.org/wiki/Fish_scale) were removed for age estimation from as many as 10 specimens from each 10 mm length interval.  All data can be obtained from `data(RockBassLO2)` from the `FSAdata` package. [*Note that the filename contains an "oh" not a "zero."*].

1. Separate the observed data into age- and length-samples.
    a. How many fish are in each sample?
```{r}
data(RockBassLO2)
rb.len <- filterD(RockBassLO2,is.na(age))
nrow(rb.len)
rb.age <- filterD(RockBassLO2,!is.na(age))
nrow(rb.age)
```

2. Add a variable to the age-sample that contains the 10 mm TL categories.
```{r}
rb.age.mod <- mutate(rb.age,lcat10=lencat(tl,w=10))
```

3. Construct a table of the **number** (not proportion) of fish in each age and 10 mm TL category in the age-sample.  From these results, compute each of the following *by hand* (i.e., not using R).
```{r}
freq <- xtabs(~lcat10+age,rb.age.mod)
```
    a. How many Rock Bass in the age-sample are in the 180 mm TL category?
```{r}
sum(freq["180",])
```
    b. How many age-7 Rock Bass are in the age-sample?
```{r}
sum(freq[,"7"])
```
    c. What proportion of Rock Bass in the 210 mm TL category are age 5?
```{r}
freq["210","5"]/sum(freq["210",])
```
    d. What proportion of Rock Bass in the 180 mm TL category are age 5?
```{r}
freq["180","5"]/sum(freq["180",])
```

4. Construct an **observed** age-length key from the table above (using R).  From these results answer the following questions.
```{r}
ALK1 <- prop.table(freq,margin=1)
```
    a. What proportion of Rock Bass in the 210 mm TL category should be assigned age 5?
```{r}
ALK1["210","5"]
```
    b. How many of thirty Rock Bass in the 180 mm TL category should be assigned age 5?
```{r}
30*ALK1["180","5"]
```
    c. Construct a plot of the **observed** age-length key.  Are there any potential anomalies in the plot that would suggest that a smoothed age-length key should be used instead?
    
```{r}
alkPlot(ALK1)
```

1. Construct a **smoothed** age-length key.  From these results answer the following questions.
```{r}
mlr <- multinom(age~lcat10,data=rb.age.mod,maxit=500)
lens <- seq(110,270,10)
ALK2 <- predict(mlr,data.frame(lcat10=lens),type="probs")
row.names(ALK2) <- lens
```
    a. What proportion of Rock Bass in the 210 mm length category should be assigned age 5?
```{r}
ALK2["210","5"]
```
    b. How many of thirty Rock Bass in the 180 mm length category should be assigned age 5?
```{r}
30*ALK2["180","5"]
```

\vspace{24pt}
5. **Make sure to save a cleaned (no code that is not needed or results in an error) script!**

----

```{r echo=FALSE, results="hide", message=FALSE}
purl2("ALK1.Rmd",moreItems=c("knitr","showkey"),topnotes="Nunavut 6-9-Aug-16")
```
