---
title: "ALK Application - Exercise"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot[RO,RE]{Ogle (Nunavut, 6-9 Aug 2016)}
- \fancyfoot[LE,LO]{ALK Application Exercise}
output: pdf_document
geometry: left=0.5in, right=0.5in, top=0.5in, bottom=0.75in
---

```{r echo=FALSE, results='hide', message=FALSE}
source("../knitr_setup.R")
showkey <- FALSE
opts_chunk$set(echo=showkey,
               results=ifelse(showkey,"markup","hide"),
               fig.show=ifelse(showkey,"asis","hide"))
```
```{r message=FALSE, warning=FALSE}
source("ALK1.R")
```

Continue with the data.frames and age-length keys constructed in the previous exercise.  You may either extend your previous script or create a new script that `source()`s the previous script.

1. Use the semi-random age assignment technique from Isermann and Knight (2005) and the **observed** age-length key to assign ages to the unaged fish in the length-sample.  Combine the age-sample and the age-assigned length-sample into a single data frame, add a variable to this data.frame that contains the 10 mm TL categories, and use the combined data frame to answer the following questions.
```{r warning=FALSE}
rb.len.mod <- alkIndivAge(ALK1,age~tl,data=rb.len)
rb.comb <- rbind(rb.age,rb.len.mod)
rb.comb <- mutate(rb.comb,lcat10=lencat(tl,w=10))
agefreq <- xtabs(~age,data=rb.comb)
lenfreq <- xtabs(~lcat10,data=rb.comb)
lensum <- Summarize(tl~age,data=rb.comb)
```
    a. How many fish are estimated to be age 5?
```{r}
agefreq[["5"]]
```
    b. How many fish are estimated to be age 11?
```{r}
agefreq[["11"]]
```
    c. Plot the age distribution for all fish.
```{r}
hist(~age,data=rb.comb,w=1,xlab="Age")
```
    d. How many fish are in the 150 mm TL interval?
```{r}
lenfreq[["150"]]
```
    e. What is the mean TL of age-5 fish?
```{r}
lensum$mean[lensum$age=="5"]
```
    f. Plot the length-at-age with the mean length-at-age superimposed for all fish.
```{r}
plot(tl~age,data=rb.comb,ylab="Total Length (mm)",xlab="Age (yrs)",
     pch=19,col=col2rgbt("black",1/20))
lines(mean~fact2num(age),data=lensum,col="blue",lwd=2)
```
    g. Compare your results from questions a-f to someone else's results (or repeat those questions).  Did you both get the *exact* same results?  Why or why not?  If not, how different were they?
\vspace{24pt}
2. [**Time Permitting**] Repeat the previous question but using the **smoothed** age-length key.
\vspace{24pt}
3. [**Time Permitting**] Fit the typical VBGF to these results.
\vspace{24pt}
4. [**Time Permitting**] Compute the instantaneous mortality rate from these results.

----

```{r echo=FALSE, results="hide", message=FALSE}
purl2("ALK2.Rmd",moreItems=c("knitr","showkey"),topnotes="Nunavut 6-9-Aug-16")
```
