---
title: "Very Quick R Basics"
subtitle: "Exercise"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot[RO,RE]{Ogle (Nunavut, 6-9 Aug 2016)}
- \fancyfoot[LE,LO]{R Basics Exercise}
output: pdf_document
geometry: left=0.5in, right=0.5in, top=0.5in, bottom=0.75in
---

```{r echo=FALSE, results='hide'}
showkey <- FALSE
source("../knitr_setup.R")
opts_chunk$set(echo=showkey, results=ifelse(showkey,"markup","hide"))
```

----

Answer the following questions with R code by creating (*and editing if you make a mistake*) an R script and iteratively running the code in RStudio.

1. Load the `FSA` library (*even though it is not required for this exercise*).
```{r message=FALSE}
library(FSA)
```

2. Use an expression to compute the CPE (number of fish per net) if 87 fish were captured in three nets.
```{r}
87/3
```

3. Assign to an object (and then view) the result of an expression that calculates the CPE (number of fish per 1000 m of gillnet per night) if 87 fish were captured in 1400 m of gillnet fished for two nights.
```{r}
( res <- 87/1400/2*1000 )
```

4. Create an expression that uses the object from the previous question to modify the CPE to be per 300 m of gillnet per night (i.e., don't re-create the previous expression, simply modify it by starting with the saved object.)
```{r}
res/1000*300
```

5. Enter the following observed catches into a vector called `ct` -- 87, 54, 12, 98, 45, 5, 78.
```{r}
ct <- c(87,54,12,98,45,5,78) 
```

6. Enter the following efforts (number of nets) into a vector called `ft` -- 3, 3, 2, 5, 2, 2, 4.
```{r}
ft <- c(3,3,2,5,2,2,4)
```

7. Use your `ct` and `ft` vectors to compute a vector called `cpe` that contains the CPE (number of fish per net).
```{r}
( cpe <- ct/ft )
```

8. Find the mean CPE.
```{r}
mean(cpe)
```

9. Use R code to find the third CPE.
```{r}
cpe[3]
```

10. Use R code to simultaneously find the third and fifth CPEs.
```{r}
cpe[c(3,5)]
```

11. Use R code to eliminate the seventh CPE (but retain the other six CPEs).
```{r}
cpe[-7]
```

12. Use R code to find the CPEs for ONLY the days when two nets were fished.
```{r}
cpe[ft==2]
```

13. Find the mean CPE for ONLY those days when three or more nets were fished.
```{r}
cpe.gt3 <- cpe[ft>=3]
mean(cpe.gt3)
```

14. Save your R script, close RStudio, open RStudio, and re-run your script.

```{r echo=FALSE, results="hide", message=FALSE}
purl2("Basics.Rmd",moreItems=c("knitr","showkey"),topnotes="Nunavut 6-9-Aug-16")
```
