---
title: "One-Way ANOVA - Nunavut Exercise"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot[RO,RE]{Ogle (Nunavut, 6-9 Aug 2016)}
- \fancyfoot[LE,LO]{One-Way ANOVA Exercise}
output: pdf_document
geometry: left=0.5in, right=0.5in, top=0.5in, bottom=0.75in
---

```{r echo=FALSE, results='hide', message=FALSE}
showkey <- FALSE
source("../knitr_Ex_setup.R")
```
```{r message=FALSE, warning=FALSE}
library(car)
library(multcomp)
library(FSA)
library(dplyr)
library(readxl)
library(plotrix)
```

----

1. Load the `PG008_original.xlsx` file into a data.frame object, create a factor version of the `year` variable, and determine if the mean fork length of fish sampled from **"Seawater"** differs among years.  [*Note, be careful about the inconsistent use of levels in `water.type`*.]
```{r message=FALSE, warning=FALSE}
source("PG008_readdata.R")
dNU_SW <- filterD(dNU,water.type %in% c("Seawater","seawater"))
```
```{r fig.width=6}
aov1 <- lm(fl~fyear,data=dNU_SW)
residPlot(aov1)
leveneTest(aov1)
anova(aov1)
mc1 <- glht(aov1,mcp(fyear="Tukey"))
summary(mc1)
cld(mc1)
```
```{r}
sum <- Summarize(fl~fyear,data=dNU_SW)
sum <- select(sum,fyear,n,mean,sd)
sum <- mutate(sum,year=fact2num(fyear),se=sd/sqrt(n),
                   LCI=mean-qt(0.975,df=n-1)*se,UCI=mean+qt(0.975,df=n-1)*se)
with(sum,plotCI(year,mean,li=LCI,ui=UCI,pch=19,xlim=c(2005.8,2012.2),ylim=c(590,790),
                     xlab="Year",ylab="Mean Fork Length (mm)"))
with(sum,text(year,UCI,c("a","b","b","b","b"),pos=3))
```

2. [*Time permitting*]  Determine if the mean fork length of **Age-15 to 20** (inclusive) Arctic Char captured in **"freshwater"** differs among years.
```{r message=FALSE, warning=FALSE}
dNU_FW <- filterD(dNU,water.type %in% c("Freshwater","freshwater"),age>=15,age<=20)
```
```{r fig.width=6}
aov2 <- lm(fl~fyear,data=dNU_FW)
residPlot(aov2)
leveneTest(aov2)
anova(aov2)
mc2 <- glht(aov2,mcp(fyear="Tukey"))
summary(mc2)
cld(mc2)
```
```{r}
sum <- Summarize(fl~fyear,data=dNU_FW)
sum <- select(sum,fyear,n,mean,sd)
sum <- mutate(sum,year=fact2num(fyear),se=sd/sqrt(n),
                   LCI=mean-qt(0.975,df=n-1)*se,UCI=mean+qt(0.975,df=n-1)*se)
with(sum,plotCI(year,mean,li=LCI,ui=UCI,pch=19,xlim=c(2001.8,2010.2),ylim=c(640,760),
                     xlab="Year",ylab="Mean Fork Length (mm)"))
with(sum,text(year,UCI,c("a","b","b"),pos=3))
```

```{r echo=FALSE, results="hide", message=FALSE}
purl2("ANOVA1.Rmd",moreItems=c("knitr","showkey"),topnotes="Nunavut 6-9-Aug-16")
```
