---
title: "von Bertalanffy Growth Function"
subtitle: "Exercise - Nunavut"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead{}
- \fancyfoot[RO,RE]{Ogle (Nunavut, 6-9 Aug 2016)}
- \fancyfoot[LE,LO]{von Bertalanffy Growth Function Exercise}
output: pdf_document
geometry: left=0.5in, right=0.5in, top=0.5in, bottom=0.75in
---

```{r echo=FALSE, results='hide', message=FALSE}
showkey <- FALSE
source("../knitr_Ex_setup.R")
```
```{r message=FALSE, warning=FALSE}
library(FSA)      # for filterD(), headtail(), col2rgbt(), vbFuns(), vbStart()
library(dplyr)    # for mutate(), select()
library(nlstools) # for nlsBoot()
library(readxl)
```

----

Answer the following questions by creating an R script and iteratively running the code in RStudio.

1. Load the `PG008_original.xlsx` file into a data.frame object and restrict the data to only those fish captured in **freshwater** in **2010**.  Use these data for the following questions.
```{r message=FALSE, warning=FALSE}
source("PG008_readdata.R")
dNU_FW <- filterD(dNU,!is.na(FL),!is.na(age),water.type %in% c("freshwater","Freshwater"),year==2010)
# Removed some obvious outliers
dNU_FW <- filterD(dNU_FW,!(age>10 & FL<200))
```

    a. Examine the plot of length versus age.  Do the data look linear or curved, is there an obvious asymptote, are young fish well represented, how variable are lengths within ages?
```{r fig.width=6}
xlbl <- "Age (yrs)"
ylbl <- "Fork Length (mm)"
clr1 <- "black"
clr2 <- col2rgbt(clr1,1/9)

plot(FL~age,data=dNU_FW,pch=19,col=clr2,xlab=xlbl,ylab=ylbl)
```

    b. Fit the typical parameterization of the VBGF to these data.  Construct a fitted-line plot (i.e., superimpose the fitted VBGF onto the length versus age plot) and a residual plot and comment on model fit.
```{r fig.show=ifelse(showkey,'hold','hide')}
vb <- vbFuns("Typical")
svb <- vbStarts(FL~age,data=dNU_FW,type="Typical",plot=TRUE)
fit1 <- nls(FL~vb(age,Linf,K,t0),data=dNU_FW,start=svb)
cf <- coef(fit1)
plot(FL~age,data=dNU_FW,xlab=xlbl,ylab=ylbl,pch=19,col=clr2)
curve(vb(x,cf),from=6,to=29,n=500,lwd=2,col=clr1,add=TRUE)
```
```{r fig.width=6, fig.show=ifelse(showkey,'hold','hide'), warning=FALSE}
residPlot(fit1)
```

    c. Compute the correlation between parameter values.  Comment
```{r}
summary(fit1,correlation=TRUE)
```

    d. Construct a table of parameter estimates with 95% bootstrap confidence intervals.  Carefully interpret the value of each parameter.  Comment on how realistic you think each estimate is and the relative widths of the confidence intervals.
```{r}
boot1 <- nlsBoot(fit1,niter=1000)
cbind(Ests=cf,confint(boot1))
```

    e. Predict the mean length, with 95% confidence interval, for an age-18 Charr.  Comment on the width of this confidence interval? 
```{r}
ageX <- 18
predict(fit1,data.frame(age=ageX))
pv <- apply(boot1$coefboot,MARGIN=1,FUN=vb,t=ageX)
quantile(pv,c(0.025,0.975))
```

----

2. [*Time Permitting*] Repeat the previous question but using either the original, Gallucci and Quinn, or Mooij parameterizations of the VBGF.  [Note that you can see the equations for these VBGFs with, for example, `growthFunShow("vonBertalanffy",param="GallucciQuinn",plot=TRUE)`.  You can declare a function for these VBGFs by using, for example, `vb <- vbFuns("GallucciQuinn")`.]  How does the fit of this model (and estimates of the common parameters) compare with the results from the typical VBGF fit in the previous question?
```{r}
vbgq <- vbFuns("GallucciQuinn")
svbgq <- vbStarts(FL~age,data=dNU_FW,type="GallucciQuinn")
fitgq <- nls(FL~vbgq(age,omega,K,t0),data=dNU_FW,start=svbgq)
cfgq <- coef(fitgq)
plot(FL~age,data=dNU_FW,xlab=xlbl,ylab=ylbl,pch=19,col=clr2)
curve(vb(x,cf),from=6,to=35,n=500,lwd=5,col=clr1,add=TRUE)
curve(vbgq(x,cfgq),from=6,to=35,n=500,lwd=2,col="red",add=TRUE)
cf
cfgq
```

----

```{r echo=FALSE, results="hide", message=FALSE}
purl2("VBGF1.Rmd",moreItems=c("knitr","showkey","seed"),topnotes="Nunavut 6-9-Aug-16")
```
